# Reproducible Research: Peer Assessment 1

Hi! Thanks for taking the time to review this assessment! We will require
the `ggplot2` and `data.table` libraries to knit this R Markdown document

```{r, echo=TRUE}
library(ggplot2)
library(data.table)
```

## Loading and preprocessing the data

First of all, we will extract the ZIP file with the activity date and load it up
on a data.frame object. We will pre-process and extract the step counts and intervals
as `numeric` and the date information as a `Date` object to make computation easier  

```{r, echo=TRUE}
df.activity = read.csv(unz("activity.zip", "activity.csv"), header=T,
                    sep=",", stringsAsFactors=F, na.strings="NA",
                    colClasses=c("numeric", "Date", "numeric"))

dt.activity = data.table(df.activity)
dt.activity
```

## What is mean total number of steps taken per day?

First of all, we need to combine all the steps that happened on each day
```{r, echo=TRUE}
dt.totalsteps = dt.activity[, sum(steps, na.rm=T), by="date"]
head(dt.totalsteps)
```

We can then create an histogram with the total steps data, and calculate their *mean* 
and *median* values, while ignoring missing values of the data

```{r, echo=TRUE}
hist_plot = qplot(dt.totalsteps$V1, geom="histogram", binwidth=3000,
            main="Total Number of Steps per Day", xlab="Steps", fill=..count..)

hist_plot

steps.median = median(dt.totalsteps$V1)
steps.median

steps.mean = mean(dt.totalsteps$V1)
steps.mean
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
