---
output:
  html_document: default
---
# Reproducible Research: Peer Assessment 1

Hi! Thanks for taking the time to review this assessment! We will require
the `ggplot2` and `data.table` libraries to knit this R Markdown document

```{r, echo=TRUE}
library(ggplot2)
library(data.table)
```

## Loading and preprocessing the data

First of all, we will extract the ZIP file with the activity date and load it up
on a data.frame object. We will pre-process and extract the step counts and intervals
as `numeric` and the date information as a `Date` object to make computation easier  

```{r, echo=TRUE}
df.activity = read.csv(unz("activity.zip", "activity.csv"), header=T,
                    sep=",", stringsAsFactors=F, na.strings="NA",
                    colClasses=c("numeric", "Date", "numeric"))

dt.activity = data.table(df.activity)
rm(df.activity)
dt.activity
```

## What is mean total number of steps taken per day?

First of all, we need to combine all the steps that happened on each day

```{r, echo=TRUE}
dt.totalsteps = dt.activity[, sum(steps, na.rm=T), by="date"]
head(dt.totalsteps)
```

We can then create an histogram with the total steps data, and calculate their *mean* 
and *median* values, while ignoring missing values of the data

```{r, echo=TRUE}
qplot(dt.totalsteps$V1, geom="histogram", binwidth=3000,
      main="Total Number of Steps per Day", xlab="Steps", fill=..count..)
median(dt.totalsteps$V1)
mean(dt.totalsteps$V1)
```

## What is the average daily activity pattern?

This time, instead of combining per day, we will combine the steps by each interval
that we have available. on the dataset.

```{r, echo=TRUE}
dt.pattern = dt.activity[, sum(steps, na.rm=T)/.N, by="interval"]
head(dt.pattern)
```

Now, we create a plot to help us visualize the pattern of daily activity averaged
over all the days.

```{r, echo=TRUE}
qplot(interval, V1, data = dt.pattern, xlab="Interval", ylab="Number of Steps",
       color=I("red"),  geom = "line", main="Daily Activity Pattern")
```

We can also figure out which interval is the one that contains the maximum number
of stepts across all the days.

```{r, echo=TRUE}
dt.pattern[V1 == max(dt.pattern$V1)]$interval
```

## Imputing missing values

First, let us verify what is the number of missing values in the dataset.

```{r, echo=TRUE}
sum(is.na(dt.activity$steps))
```

We will fill these missing vaules by using the mean of the specific 5-minute interval
as a replacement for the missing value. I already have these quantities calculated
in the `dt.pattern` table.

```{r, echo=TRUE}
dt.nomissing = copy(dt.activity)
dt.nomissing[is.na(steps), steps := dt.pattern[dt.pattern$interval == interval]$V1]
sum(is.na(dt.nomissing$steps))
```

Now, we again create an histogram with the total steps data, and calculate their *mean* 
and *median* values to verify the changes we made on the dataset.

```{r, echo=TRUE}
dt.totalsteps.noNA = dt.nomissing[, sum(steps, na.rm=T), by="date"]
head(dt.totalsteps.noNA)
qplot(dt.totalsteps.noNA$V1, geom="histogram", binwidth=3000,
      main="Total Number of Steps per Day (no NAs)", xlab="Steps", fill=..count..)
median(dt.totalsteps.noNA$V1)
mean(dt.totalsteps.noNA$V1)
```

As a result from this change, the mean got really close to the median, and the
shape of the histogram got much closer to a bell shape. The concentration of
days with less then 3000 steps got really small, as most of them were filled with
missing values, and that went to the bucket that contains the mean and median.

## Are there differences in activity patterns between weekdays and weekends?

First, let us separate our dates in weekdays and weekends so we can start splitting
the interval counts among them. For this we will create a factor variable and
add it to the dataset

```{r, echo=TRUE}
activity.weekdays = weekdays(dt.activity$date)
factor.weekday = factor( ifelse(activity.weekdays == "Saturday" 
                                | activity.weekdays == "Sunday",
                                "weekend",
                                "weekday"),
                         levels = c("weekend","weekday"))
dt.activity[, weekstatus := factor.weekday]
```

We then proceed to do some aggregating magic with our data.table

```{r, echo=TRUE}
dt.pattern.week = dt.activity[, sum(steps, na.rm=T)/.N, by=c("weekstatus", "interval")]
head(dt.pattern.week)
```

Now, we create a plot to help us visualize the pattern of daily activity averaged
over weekends and weekdays.

```{r, echo=TRUE}
qplot(interval, V1, data = dt.pattern.week, facets = weekstatus ~ .,
      xlab="Interval", ylab="Number of Steps",
      color=I("red"),  geom = "line", main="Daily Activity Pattern")
``````
