---
output:
  html_document: default
---
# Reproducible Research: Peer Assessment 1

Hi! Thanks for taking the time to review this assessment! We will require
the `ggplot2` and `data.table` libraries to knit this R Markdown document

```{r, echo=TRUE}
library(ggplot2)
library(data.table)
```

## Loading and preprocessing the data

First of all, we will extract the ZIP file with the activity date and load it up
on a data.frame object. We will pre-process and extract the step counts and intervals
as `numeric` and the date information as a `Date` object to make computation easier  

```{r, echo=TRUE}
df.activity = read.csv(unz("activity.zip", "activity.csv"), header=T,
                    sep=",", stringsAsFactors=F, na.strings="NA",
                    colClasses=c("numeric", "Date", "numeric"))

dt.activity = data.table(df.activity)
rm(df.activity)
dt.activity
```

## What is mean total number of steps taken per day?

First of all, we need to combine all the steps that happened on each day

```{r, echo=TRUE}
dt.totalsteps = dt.activity[, sum(steps, na.rm=T), by="date"]
head(dt.totalsteps)
```

We can then create an histogram with the total steps data, and calculate their *mean* 
and *median* values, while ignoring missing values of the data

```{r, echo=TRUE}
qplot(dt.totalsteps$V1, geom="histogram", binwidth=3000,
      main="Total Number of Steps per Day", xlab="Steps", fill=..count..)
median(dt.totalsteps$V1)
mean(dt.totalsteps$V1)
```

## What is the average daily activity pattern?

This time, instead of combining per day, we will combine the steps by each interval
that we have available. on the dataset.

```{r, echo=TRUE}
dt.pattern = dt.activity[, sum(steps, na.rm=T), by="interval"]
head(dt.pattern)
```

Now, we create a plot to help us visualize the pattern of daily activity averaged
over all the days.

```{r, echo=TRUE}
qplot(interval, V1, data = dt.pattern, xlab="Interval", ylab="Number of Steps",
       color=I("red"),  geom = "line", main="Daily Activity Pattern")
```

We can also figure out which interval is the one that contains the maximum number
of stepts across all the days.

```{r, echo=TRUE}
dt.pattern[V1 == max(dt.pattern$V1)]$interval
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
